ms2

### 2017-05-29 ms2/ms2.Rmd
ms2 year average

@@

# Consumption data

@@

## Aggregate usage correlates with fraction of users

## The two inequality metrics not correlate well

```{r nzgini_mean}
all_ineq %>%
  ggplot(aes(x=mean_fnz, xmin=mean_fnz-sem_fnz, xmax=mean_fnz+sem_fnz,
             y=mean_nzgini, ymin=mean_nzgini-sem_nzgini, ymax=mean_nzgini+sem_nzgini,
             color=region, label=state_abbreviation)) +
  geom_point() +
  geom_linerange() +
  geom_errorbarh(height=0) +
  geom_text(nudge_x=0.04) +
  xlab('mean fraction with at least one claim') +
  ylab('mean Gini coefficient') +
  theme_minimal()
```

# Resistance data

@@

And I join the resistance drugs into classes:

@@

@@

Significant results using Spearman correlations of mean percent nonsusceptible ~ claims per beneficiary:

```{r spearman_models, cache=FALSE}
results %>%
  filter(model=='spearman') %>%
  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
  filter(p.value < 0.05) %>%
  arrange(p.adj) %>%
  select(bug, drug_group, estimate, p.value, p.adj, n) %>%
  kable
```

```{r}
interval_plot = function(df) {
  arrange(df, estimate) %>%
  mutate(bug_drug=interaction(bug, drug_group),
         bug_drug=factor(bug_drug, levels=bug_drug, ordered=TRUE)) %>% 
    ggplot(aes(x=bug_drug, y=estimate, ymin=ci_low, ymax=ci_high, color=drug_group)) +
    geom_point() +
    geom_linerange() +
    coord_flip()
}

results %>%
  filter(model=='spearman') %>%
  interval_plot()
```

@@

And the same thing using HRRs:

@@

## Univariable

The model here is mean percent nonsusceptible ~ claims per beneficiary. The significant results are:

@@

And the same thing with HRRs:

@@

## Multivariable

The model here is mean percent nonsusceptible ~ fraction nonzero + nonzero Gini. In some of the models, the nonzero Gini term was significant, but not after adjusting for multiple hypotheses. I show just those terms that met the unadjusted threshold:

```{r}
results %>%
  filter(model=='multi', term==multi_term) %>%
  interval_plot()
```

```{r}
results %>%
  filter(model=='multi', term==multi_term) %>%
  mutate(bug_drug=interaction(bug, drug_group)) %>%
  ggplot(aes(x=p.value, y=estimate, color=drug_group)) +
  geom_point() +
  scale_color_brewer(palette='Set1')
```

And what about looking at multivariate only for ones that passed the threshold for univariable?

```{r eval=FALSE}
sig_bug_drug = results %>%
  filter(model=='uni', term==uni_term) %>%
  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
  filter(p.adj < 0.05) %>%
  select(bug, drug_group)

results %>%
  filter(model=='multi', term==multi_term) %>%
  right_join(sig_bug_drug, by=c('bug', 'drug_group')) %>%
  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
  select(bug, drug_group, estimate, p.value, p.adj) %>%
  kable
```

Using HRRs, adjusting $p$ with all possibilities:

And then doing multiple hypothesis correction only for those that made it through the first model:

```{r hrr_multi_models}
sig_bug_drug = hrr_results %>%
  filter(model=='uni', term==uni_term) %>%
  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
  filter(p.adj < 0.05) %>%
  select(bug, drug_group)

hrr_results %>%
  filter(model=='multi', term==multi_term) %>%
  right_join(sig_bug_drug, by=c('bug', 'drug_group')) %>%
  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
  arrange(p.value) %>%
  select(bug, drug_group, estimate, p.value, p.adj) %>%
  kable
```

## Two metrics

It turns out that the different drugs have different relationships with the two metrics. Each point shows the consumption characteristics for a drug group in a state.

```{r ecoli_quin, eval=FALSE}
left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
  left_join(regions, by='state') %>%
  filter(bug=='E. coli', drug_group=='quinolone') %>%
  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, color=region)) +
  geom_smooth(aes(x=mean, y=mean_percent_nonsusceptible, color=NULL), method='lm', col='black') +
  geom_point() +
  xlab('quinolone claims per beneficiary') + ylab('mean percent nonsusceptible E. coli isolates') +
  theme_minimal()

left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
  left_join(regions, by='state') %>%
  filter(bug=='E. coli', drug_group=='quinolone') %>%
  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, label=state)) +
  geom_smooth(method='lm', col='black', se=FALSE) +
  geom_point() +
  geom_text() +
  xlim(0.2, 0.45) + ylim(0, 50) +
  xlab('quinolone claims per beneficiary') + ylab('mean percent nonsusceptible E. coli isolates') +
  theme_minimal()
```

Bactrim resistance in *E. coli*. First, the aggregate relationship, then the relationship between "excess" resistance (the residual of resistance against aggregate consumption) against nonzero Gini.

```{r ecoli_bactrim, eval=FALSE}
ecoli_bactrim_dat = left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
  left_join(regions, by='state') %>%
  filter(bug=='E. coli', drug_group=='tmp_smx') #%>%
  #filter(!is.na(mean_percent_nonsusceptible))

ecoli_bactrim_dat %>%
  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, label=state)) +
  geom_smooth(method='lm', col='gray', se=FALSE) +
  geom_point() +
  geom_text() +
  scale_x_continuous(limits=c(0.08, 0.2), breaks=c(0.08, 0.1, 0.15, 0.2)) +
  ylim(10, 45) +
  xlab('TMP/SMX claims per beneficiary') + ylab('mean percent nonsusceptible E. coli isolates') +
  theme_minimal()

ecoli_bactrim_dat %>%
  { mutate(., excess_resistance=residuals(lm(mean_percent_nonsusceptible ~ mean, data=.))) } %>%
  ggplot(aes(x=nzgini, y=excess_resistance)) +
  geom_smooth(method='lm', col='gray', se=FALSE) +
  geom_point() +
  scale_x_continuous(limits=c(0.25, 0.375)) +
  scale_y_continuous(limits=c(-12, 20), breaks=c(-12, 0, 10, 20)) +
  xlab('Skew towards high consumers (Gini coefficient)') +
  ylab('"Excess" E. coli resistance (see caption)') +
  theme_minimal()

ecoli_bactrim_dat %>%
  ggplot(aes(x=nzgini, y=mean_percent_nonsusceptible, label=state)) +
  geom_smooth(method='lm', col='gray', se=FALSE) +
  geom_point() +
  geom_text() +
  xlab('Gini coefficient)') +
  ylab('E. coli resistance') +
  theme_minimal()
```

```{r, eval=FALSE}
left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
  left_join(regions, by='state') %>%
  filter(bug=='S. pneumoniae', drug_group=='macrolide') %>%
  filter(!is.na(mean_percent_nonsusceptible)) %>% 
  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, label=state)) +
  geom_smooth(method='lm', col='gray', se=FALSE) +
  geom_point() +
  geom_text() +
  theme_minimal()
```

```{r, eval=FALSE}
left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
  left_join(regions, by='state') %>%
  filter(bug=='Eb. cloacae', drug_group=='quinolone') %>%
  filter(!is.na(mean_percent_nonsusceptible)) %>%
  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, label=state)) +
  geom_smooth(method='lm', col='gray', se=FALSE) +
  geom_point() +
  geom_text() +
  theme_minimal()
```

```{r mssa_clinda, eval=FALSE}
dat = left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
  left_join(regions, by='state') %>%
  filter(bug=='MSSA', drug_group=='clindamycin') #%>%
  #filter(!is.na(mean_percent_nonsusceptible))

dat %>%
  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, label=state)) +
  geom_smooth(method='lm', col='gray', se=FALSE) +
  geom_point() +
  geom_text() +
  #scale_x_continuous(limits=c(0.08, 0.2), breaks=c(0.08, 0.1, 0.15, 0.2)) +
  #ylim(10, 45) +
  xlab('clindamycin claims per beneficiary') + ylab('mean percent nonsusceptible MSSA isolates') +
  theme_minimal()

dat %>%
  { mutate(., excess_resistance=residuals(lm(mean_percent_nonsusceptible ~ mean, data=.))) } %>%
  ggplot(aes(x=nzgini, y=excess_resistance)) +
  geom_smooth(method='lm', col='gray', se=FALSE) +
  geom_point() +
  #scale_x_continuous(limits=c(0.25, 0.375)) +
  #scale_y_continuous(limits=c(-12, 20), breaks=c(-12, 0, 10, 20)) +
  xlab('Gini coefficient') +
  ylab('"Excess" MSSA resistance') +
  theme_minimal()
```

## MRSA clinda

```{r, eval=FALSE}
left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
  left_join(regions, by='state') %>%
  filter(bug=='MRSA', drug_group=='clindamycin') %>%
  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, label=state)) +
  geom_smooth(method='lm', col='gray', se=FALSE) +
  geom_point() +
  geom_text() +
  #scale_x_continuous(limits=c(0.08, 0.2), breaks=c(0.08, 0.1, 0.15, 0.2)) +
  #ylim(10, 45) +
  xlab('clindamycin claims per beneficiary') + ylab('mean percent nonsusceptible MRSA isolates') +
  theme_minimal()
```

>>>>

# To do

- filter by HRRs, not states

@@

@@

@@

@@

# Landscape of bug/drug correlations

## Selecting correlations to test

I join the resistance drugs into classes:

## Preparing the consumption data

Unlike previous analyses, this time I treated the consumption and resistance data similarly: I averaged both across all data years. Thus, to get a state's Gini coefficient, I computed that number each year and then took the average.

@@

Significant results using Spearman correlations of mean percent nonsusceptible ~ claims per beneficiary. A star after the bug/drug combination implies that the adjusted $p < 0.05$, and a dot means the unadjusted $p < 0.05$.

```{r spearman_models}
results %>%
  filter(model=='spearman') %>%
  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
  filter(p.value < 0.05) %>%
  arrange(p.adj) %>%
  select(bug, drug_group, estimate, p.value, p.adj, n) %>%
  kable
```

Aside from individual significant results, we can look at the landscape of correlations. They are almost all positive, with MRSA/Bactrim making an interesting exception. Lines show 95% confidence intervals.

```{r spearman_interval}
interval_plot = function(df) {
  arrange(df, estimate) %>%
  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
  mutate(sig=case_when(.$p.adj < 0.05 ~ '*',
                       .$p.value < 0.05 ~ '.',
                       TRUE ~ ''),
         bug_drug=interaction(bug, drug_group, sig),
         bug_drug=factor(bug_drug, levels=bug_drug, ordered=TRUE)) %>% 
    ggplot(aes(x=bug_drug, y=estimate, ymin=ci_low, ymax=ci_high, color=drug_group)) +
    geom_point() +
    geom_linerange() +
    coord_flip()
}

results %>%
  filter(model=='spearman') %>%
  interval_plot() +
  ylab("Spearman's rho")
```

## Differences in consumption/resistance patterns explain some of these results

Different bug/drug combinations have different spreads along the consumption and resistance axes, so we would expect to see different strengths of association for each one.

# Unevenness of consumption and antibiotic resistance

## Drugs are consumed differently in different places

## Effect of unevenness on antibiotic resistance

Here we use the same bug/drug combinations, but instead we test the multivariable model mean percent nonsusceptible ~ fraction nonzero + nonzero Gini.

There is only one significant and one marginally significant result:

```{r}
results %>%
  filter(model=='multi', term==multi_term) %>%
  interval_plot() +
  ylab('Gini coefficient')
```

Using HRRs, one of the answers remains the same, but a different marginally-significant result pops up.

```{r}
hrr_results %>%
  filter(model=='multi', term==multi_term) %>%
  interval_plot()
```

```{r ecoli_quin}
left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
  left_join(regions, by='state') %>%
  filter(bug=='E. coli', drug_group=='quinolone') %>%
  ggplot(aes(x=cpb, y=mean_percent_nonsusceptible, label=state)) +
  geom_smooth(method='lm', col='gray', se=FALSE) +
  geom_point() +
  geom_text() +
  xlim(0.2, 0.45) + ylim(0, 50) +
  xlab('quinolone claims per beneficiary') + ylab('mean percent nonsusceptible E. coli isolates') +
  theme_minimal()
```

```{r ecoli_bactrim}
ecoli_bactrim_dat = left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
  left_join(regions, by='state') %>%
  filter(bug=='E. coli', drug_group=='tmp_smx') #%>%
  #filter(!is.na(mean_percent_nonsusceptible))

ecoli_bactrim_dat %>%
  ggplot(aes(x=cpb, y=mean_percent_nonsusceptible, label=state)) +
  geom_smooth(method='lm', col='gray', se=FALSE) +
  geom_point() +
  geom_text() +
  scale_x_continuous(limits=c(0.08, 0.2), breaks=c(0.08, 0.1, 0.15, 0.2)) +
  ylim(10, 45) +
  xlab('TMP/SMX claims per beneficiary') + ylab('mean percent nonsusceptible E. coli isolates') +
  theme_minimal()
```

```{r}
ecoli_bactrim_dat %>%
  ggplot(aes(x=nzg, y=mean_percent_nonsusceptible, label=state)) +
  geom_smooth(method='lm', col='gray', se=FALSE) +
  geom_point() +
  geom_text() +
  xlab('Gini coefficient)') +
  ylab('E. coli resistance') +
  theme_minimal()
```

```{r}
left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
  left_join(regions, by='state') %>%
  filter(bug=='S. pneumoniae', drug_group=='macrolide') %>%
  filter(!is.na(mean_percent_nonsusceptible)) %>% 
  ggplot(aes(x=cpb, y=mean_percent_nonsusceptible, label=state)) +
  geom_smooth(method='lm', col='gray', se=FALSE) +
  geom_point() +
  geom_text() +
  theme_minimal()
```

```{r}
left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
  left_join(regions, by='state') %>%
  filter(bug=='Eb. cloacae', drug_group=='quinolone') %>%
  filter(!is.na(mean_percent_nonsusceptible)) %>%
  ggplot(aes(x=cpb, y=mean_percent_nonsusceptible, label=state)) +
  geom_smooth(method='lm', col='gray', se=FALSE) +
  geom_point() +
  geom_text() +
  theme_minimal()
```

```{r mssa_clinda}
left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
  left_join(regions, by='state') %>%
  filter(bug=='MSSA', drug_group=='clindamycin') %>%
  ggplot(aes(x=cpb, y=mean_percent_nonsusceptible, label=state)) +
  geom_smooth(method='lm', col='gray', se=FALSE) +
  geom_point() +
  geom_text() +
  xlab('clindamycin claims per beneficiary') + ylab('mean percent nonsusceptible MSSA isolates') +
  theme_minimal()
```

# Supplementary results

## Metrics of unevenness

### Aggregate usage correlates with fraction of users

### The two inequality metrics not correlate well

```{r nzgini_mean}
all_ineq %>%
  ggplot(aes(x=mean_fnz, xmin=mean_fnz-sem_fnz, xmax=mean_fnz+sem_fnz,
             y=mean_nzgini, ymin=mean_nzgini-sem_nzgini, ymax=mean_nzgini+sem_nzgini,
             color=region, label=state_abbreviation)) +
  geom_point() +
  geom_linerange() +
  geom_errorbarh(height=0) +
  geom_text(nudge_x=0.00) +
  xlab('mean fraction with at least one claim') +
  ylab('mean Gini coefficient') +
  theme_minimal()
```

## Spearman models

### State-level volcano plot

@@

### HRR-level correlations

@@

## Univariable regressions

Like the Spearman models, the model here is mean percent nonsusceptible ~ claims per beneficiary, only this time it's a normal linear regression.

### State-level

@@

### HRR-level

@@
